---
title: "SDS315 HW8 - Palash Pawar ppp625"
author: https://github.com/palashpawar/SDS315
output: pdf_document
date: "2025-04-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

```{r, include=FALSE, echo=FALSE, message=FALSE}
library(readr)

creatinine <- read.csv("creatinine.csv")

model <- lm(creatclear ~ age, data = creatinine)

predict(model, newdata = data.frame(age = 55))

summary(model)

pred_40 <- predict(model, newdata = data.frame(age = 40))
pred_60 <- predict(model, newdata = data.frame(age = 60))

actual_40 <- 135
actual_60 <- 112

diff_40 <- actual_40 - pred_40
diff_60 <- actual_60 - pred_60

diff_40
diff_60


```

## A) Expected creatinine clearance rate for a 55-year-old

I used the linear regression equation: creatclear = 147.81 - 0.62 x age

**For a 55-year-old**: creatclear = 147.81 - 0.62 x 55 = 113.71 mL/min

This was calculated using the regression formula from the model.

## B) How does creatinine clearance rate change with age?

The slope of the regression line is: -0.62 mL/min per per. This means that for every additional year of age, the expected creatinine clearance rate decreases by approximately 0.62 mL/min.

## C) Who is healthier for their age?

I computed each person's expected rate and compared it to their actual rate:

**40-year-old:**

Expected = 147.81 - 0.62 x 40 = 123.01

Difference = 135 - 123.01 = +11.97

**60-year-old:**

Expected = 147.81 - 0.62 x 60 = 110.61

Difference = 112 - 110.61 = +1.39

Conclusion: The 40-year-old with a rate of 135 is healthier for their age, as their creatinine clearance rate is nearly 12 mL/min above the expected, compared to only 1.4 mL/min above the expected for the 60-year-old.

# Problem 2

```{r, output=FALSE, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(boot)

covid <- read_csv("covid.csv")

italy <- filter(covid, country == "Italy" & deaths > 0)
spain <- filter(covid, country == "Spain" & deaths > 0)

italy_model <- lm(log(deaths) ~ days_since_first_death, data = italy)
spain_model <- lm(log(deaths) ~ days_since_first_death, data = spain)

italy_growth <- coef(italy_model)[2]
spain_growth <- coef(spain_model)[2]

italy_doubling <- log(2) / italy_growth
spain_doubling <- log(2) / spain_growth


```

```{r, echo=FALSE}
growth_stats <- function(data, indices) {
  d <- data[indices, ]
  model <- lm(log(deaths) ~ days_since_first_death, data = d)
  growth <- coef(model)[2]
  doubling <- log(2) / growth
  return(c(growth, doubling))
}

set.seed(123)
italy_boot <- boot(data = italy, statistic = growth_stats, R = 1000)
italy_ci <- boot.ci(italy_boot, type = "perc", index = 1) # for growth
italy_dt_ci <- boot.ci(italy_boot, type = "perc", index = 2) # for doubling

set.seed(123)
spain_boot <- boot(data = spain, statistic = growth_stats, R = 1000)
spain_ci <- boot.ci(spain_boot, type = "perc", index = 1) # for growth
spain_dt_ci <- boot.ci(spain_boot, type = "perc", index = 2) # for doubling

# Print Italy results
cat("Italy Growth Rate:", round(italy_growth, 3), 
    "with 95% CI [", round(italy_ci$percent[4], 3), ",", round(italy_ci$percent[5], 3), "]\n")
cat("Italy Doubling Time:", round(italy_doubling, 1), 
    "days with 95% CI [", round(italy_dt_ci$percent[4], 1), ",", round(italy_dt_ci$percent[5], 1), "]\n\n")

# Print Spain results
cat("Spain Growth Rate:", round(spain_growth, 3), 
    "with 95% CI [", round(spain_ci$percent[4], 3), ",", round(spain_ci$percent[5], 3), "]\n")
cat("Spain Doubling Time:", round(spain_doubling, 1), 
    "days with 95% CI [", round(spain_dt_ci$percent[4], 1), ",", round(spain_dt_ci$percent[5], 1), "]\n")

```

```{r, echo=FALSE, message=FALSE}
# Combine both countries
covid_filtered <- filter(covid, deaths > 0)

ggplot(covid_filtered, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(linewidth = 1) +
  labs(
    title = "COVID-19 Daily Reported Deaths: Italy vs. Spain",
    x = "Days Since First Death",
    y = "Reported Daily Deaths"
  ) +
  theme_minimal() +
  theme(legend.title = element_blank())

```

# Problem 3

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(readr)

milk <- read_csv("milk.csv")

milk <- milk |>
  mutate(log_sales = log(sales),
         log_price = log(price))

model <- lm(log_sales ~ log_price, data = milk)
summary(model)

beta_hat <- coef(model)["log_price"]

set.seed(42)
n_iterations <- 1000
boot_coefs <- numeric(n_iterations)

for (i in 1:n_iterations) {
  sample_data <- milk[sample(nrow(milk), replace = TRUE), ]
  boot_model <- lm(log_sales ~ log_price, data = sample_data)
  boot_coefs[i] <- coef(boot_model)["log_price"]
}

ci <- quantile(boot_coefs, c(0.025, 0.975))
ci_lower <- ci[1]
ci_upper <- ci[2]

cat("Elasticity Estimate:", beta_hat, "\n")
cat("95% Bootstrapped CI: [", ci_lower, ",", ci_upper, "]\n")



```

### Conclusion:

To estimate the price elasticity of demand for milk, I log-transformed both price and sales to linearize the power-law model, then fit a linear regression of log(sales) on log(price). The slope of this regression gives the elasticity estimate of -1.62. I used bootstrapping with 1,000 resamples to compute a 95% confidence interval which was [-1.78, -1.46].
